---
title: "clean_2019"
author:
- name: Robert Day
  affiliation: Royal Perth Hospital
- name: Tegan Rourke
  affiliation: Sir Charles Gaidner Hospital
- name: Hilary
  affiliation: The Conference Company

date: "7/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r prologue, include=FALSE}
# Load needed libraries and set file locations
library(data.table)
library(stringr)
library(ggplot2)

raw_path <- "../data-raw"
clean_path <- "../data"
```

## tidy & split the submission data

Data from the submission date sheet were split into seperate files to make it easier to extend this anaylsis in the future.
All data were converted to a common format and private information removed from as needed.
```{r split merged}
# tidy & split the submission data

# Read the raw submissions to get a submission date for each abstract
MS <- fread(file.path(raw_path, "Registration.csv"))
# Filenames without spaces are safer. The Perth data is EPSM & AOCMP combined
setnames(MS, c("day_prior", "Hobart_2017", "Adelaide_2018", "Combined_2019"))
# Clean up the day_prior column, using only digits before the first " "
MS[, day_prior := lapply(tstrsplit(day_prior, " ")[1], as.integer)]
# and make them negative days, as they are days prior to conference
MS[, day_prior := -1 * abs(day_prior)]

# Write seperate files for each conference, to make future re-use easier
for (conf in names(MS)[-1]) { 
  cols  <- c("day_prior", conf)
  write.csv(na.omit(MS[, ..cols]), # only rows with values for this conf
            file = file.path(clean_path, paste0(conf, ".csv")),
            row.names = FALSE
  )
}
```

## tidy & anonymise the submission details

EPSM and AOCMP abstracts were seperated in a two stage process.
First the list of AOCMP ID numbers was used to label the abstract ID by conference.
A list of abstract dates was generated by replicating dates for days with multiple submissions.
This list was then be matched with the abstracts, which were in order of submission.
There were a few more abstract IDs than recieved submissions, all AOCMP submissions.
These were all assigned to the final date in the submissions list.

```{r munge}
# tidy & anonymise the submission details
AOCMP <- fread(file.path(raw_path, "AOCMP_abstracts.csv"))
setnames(AOCMP, "ID")
# Read the complete submission sheet
PER_19 <- fread(file.path(raw_path, "EPSM2019_subs.csv"))
setnames(PER_19, c("submitter", "city", "state", "country", "code", 
                   "theme","sub_theme", "ID"))
# Start by labelling everything as EPSM
PER_19[, conference := "EPSM"]
# Re-label rows that match the AOCMP ID list
PER_19[AOCMP, conference := "AOCMP", on = "ID"]
# Expand submission days so get a day for each submitted abstract
sub_days <- MS[!is.na(Combined_2019), rep(day_prior, Combined_2019)]
# If not enough days, assume the rest were all on the last day
all_days <- c(sub_days, rep(last(sub_days), nrow(PER_19)-length(sub_days)))
# As the abstracts are in submission order, this will set the right day
# Make the days prior negative, as they all come before conference date
PER_19[, day_prior := -1 * abs(all_days)]

# Write out seperate EPSM and AOCMP submission files (already have the combined one above)
write.csv(PER_19[conference == "EPSM", .(Perth_2019 = .N), by = day_prior],
          file = file.path(clean_path, "Perth_2019.csv"),
          row.names = FALSE
)
write.csv(PER_19[conference == "AOCMP",  .(AOCMP_2019 = .N), by = day_prior],
          file = file.path(clean_path, "AOCMP_2019.csv"),
          row.names = FALSE
)
# Write the details file, combining theme & sub-theme and dropping postcode
PER_19[, theme := paste(theme, sub_theme)]
write.csv(PER_19[,  .(submitter, city, state, country, theme, ID, conference, day_prior)],
          file = file.path(clean_path, "Perth_2019_details.csv"),
          row.names = FALSE
)
```
